Late Fall Associated Region: Genome Wide Association Study and Marker Selection

### 2025-03-26 001-map-of-samples.Rmd
Notebooks all rendering.

---
title: "Make a Map for our Paper"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('tex/citation.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

@@

```{r}
library(terra)
library(sf)
library(tidyverse)
library(ggspatial)
#library(ggsn)
library(cowplot)
library(plotly)
```

@@

# Citations

::: {#refs}
:::

>>>>

---
title: "Make a Map for our Paper"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

@@

```{r}
library(terra)
library(sf)
library(tidyverse)
library(ggspatial)
#library(ggsn)
library(cowplot)
library(plotly)
library(maps) # needed for map_data()
library(mapproj)  # making it explicit for renv
```

@@

### 2025-03-26 002-GSI_and_Fst.Rmd
Notebooks all rendering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

@@

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```

### 2025-03-26 004-lfar-and-wrap-allele-freqs.Rmd
Notebooks all rendering.

---
title: "LFAR and WRAP Allele Freqs"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('tex/citation.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

>>>>

---
title: "LFAR and WRAP Allele Freqs"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

### 2025-03-26 005-plotting-genome-locations-of-markers.Rmd
Notebooks all rendering.

---
title: "Plotting Marker Locations in Genome"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('tex/citation.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

@@

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    locus_info = "inputs/Calif-Chinook-Amplicon-Panel-Information.csv",
    chr_len = "inputs/Otsh_v2.0_chrom_lengths.tsv"
  )
  # outputs:
  output_list <- list(
    pdf = "results/genomic-locations-plot/gen-loc.pdf"
  )
}

# add the intermediate results/paths we need
output_list <- c(
  output_list,
  pdfcrop = "results/genomic-locations-plot/gen-loc-crop.pdf",
  pdftex = "tex/images/genomic-locations-plot.pdf"
)

@@

Write the thing out to a PDF file.
```{r}
ggsave(final, file = output_list$pdf, width = 7, height = 8.5)
```

Then pdfcrop it and put it where we can include it in the Supplement.
```{r}
# let's crop that down for use, too
CALL <- paste("pdfcrop ", output_list$pdf, collapse = " ")
system(CALL)

file.copy(from = output_list$pdfcrop, to = output_list$pdftex, overwrite = TRUE)
```

# Citations

::: {#refs}
:::

>>>>

---
title: "Plotting Marker Locations in Genome"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

@@

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    locus_info = "inputs/Calif-Chinook-Amplicon-Panel-Information.csv",
    chr_len = "inputs/Otsh_v2.0_chrom_lengths.tsv"
  )
  # outputs:
  output_list <- list(
    genome_locations_tex = "tex/images/genomic-locations-plot.pdf"
  )
}

# add the intermediate results/paths we need
output_list <- c(
  output_list,
  genome_locations_pdf = "results/genomic-locations-plot/gen-loc.pdf",
  genome_locations_crop = "results/genomic-locations-plot/gen-loc-crop.pdf"
)

@@

Write the thing out to a PDF file.
```{r}
ggsave(final, file = output_list$genome_locations_pdf, width = 7, height = 8.5)
```

Then pdfcrop it and put it where we can include it in the Supplement.
```{r}
# let's crop that down for use, too
CALL <- paste("pdfcrop ", output_list$genome_locations_pdf, collapse = " ")
system(CALL)

file.copy(from = output_list$genome_locations_crop, to = output_list$genome_locations_tex, overwrite = TRUE)
```

### 2025-03-26 006-winter-run-vs-all-cv-alle-freqs.Rmd
Notebooks all rendering.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

@@

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    wrap_chr_len = "inputs/wrap/chrom_lengths.txt",
    win_non_win_abs_diffs = "stored_results/win-non-win-abs-diffs-gt0.5.rds"
  )
  # outputs:
  output_list <- list(
    mh_plot = "results/wrap-discover/winter-v-non-winter-mh-plot_nind_ge10.pdf",
    mh4tex = "tex/images/winter-v-non-winter-mh-plot_nind_ge10.pdf",
    pdf = "results/genomic-locations-plot/gen-loc.pdf",
    hundy_kb_plot = "results/wrap-discover/100-kb-intervals-plot.pdf",
    hundy_kb4tex = "tex/images/wrap-slide-window.pdf",
    fuv2 = "results/wrap-discover/follow_up_variants2.rds",
    wrap61_plot = "results/wrap-discover/wrap-candi.pdf",
    wrap61_4tex = "tex/images/wrap-candi.pdf"
  )
}
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

@@

file.copy(from = output_list$mh_plot, to = output_list$mh4tex, overwrite = TRUE)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

@@

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    wrap_chr_len = "inputs/wrap/chrom_lengths.txt",
    win_non_win_abs_diffs = "stored_results/win-non-win-abs-diffs-gt0.5.rds"
  )
  # outputs:
  output_list <- list(
    winter_v_nonwinter_mh_tex = "tex/images/winter-v-non-winter-mh-plot_nind_ge10.pdf",
    hundy_kb4tex = "tex/images/wrap-slide-window.pdf",
    wrap61_4tex = "tex/images/wrap-candi.pdf"
  )
}

# now add the intermediates
output_list <- c(
  output_list,
  mh_plot = "results/wrap-discover/winter-v-non-winter-mh-plot_nind_ge10.pdf",
  hundy_kb_plot = "results/wrap-discover/100-kb-intervals-plot.pdf",
  wrap61_plot = "results/wrap-discover/wrap-candi.pdf",
  fuv2 = "results/wrap-discover/follow_up_variants2.rds"
)

# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

@@

file.copy(from = output_list$mh_plot, to = output_list$winter_v_nonwinter_mh_tex, overwrite = TRUE)
```

### 2025-03-26 007-whoa-on-chinook.Rmd
Notebooks all rendering.

---
title: "Running the California Microhap Baseline through `whoa` and Deciding We Should Drop Two of the Amplicons"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

@@

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  input_list <- list(
    final_baseline = "data/SWFSC-chinook-reference-baseline.csv.gz",
    pop_labels = "inputs/reference-collection-names.csv"
  )
  # outputs:
  output_list <- list(
    zs = "results/whoa/heterozygote-z-scores.pdf"
  )
}

@@

  ggsave(zs, filename = output_list$zs, width = 20, height = 30)
```

@@

So, what we see is that most of the really egregious ones are gone, but in some
populations there are a few that look a little wonky, but those are only
in a few pops.  So, I don't think we really need to toss any others.

>>>>

---
title: "Running the California Microhap Baseline through `whoa` and Deciding We Could Drop Two of the Amplicons in a Future Revision of the Baseline"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()

@@

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  input_list <- list(
    final_baseline = "data/SWFSC-chinook-reference-baseline.csv.gz",
    pop_labels = "inputs/reference-collection-names.csv"
  )
  # outputs:
  output_list <- list(
    whoa_zs = "results/whoa/heterozygote-z-scores.pdf"
  )
}

@@

  ggsave(zs, filename = output_list$whoa_zs, width = 20, height = 30)
```

@@

So, what we see is that most of the really egregious ones are gone, but in some
populations there are a few that look a little wonky, but those are only
in a few pops.  So, I don't think we really would be called upon to toss any others
in a future revision of the baseline.

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```

### 2025-03-26 008-lfar-gwas-and-markers.Rmd
Notebooks all rendering.

@@

```{r}
start_time <- Sys.time()

# inputs:
input_list <- list(
  chr34_late_fall_mafs = "stored_results/5Mb_Chr34-late-fall.mafs.gz",
  chr34_cv_fall_mafs = "stored_results/5Mb_Chr34-cv-fall.mafs.gz",
  gwas_lf_v_sanjo = "stored_results/lfar-gwas/lf_v_sanjo-NC_037130.1.lrt0.gz",
  gwas_lf_v_frhf = "stored_results/lfar-gwas/lf_v_frhf-NC_037130.1.lrt0.gz",
  ots34_snpeff = "stored_results/thompson2020-vcf/ots34-5Mb-snpEff.vcf.gz",
  pvals = "stored_results/stored_p_vals.rds"
)
# outputs:
output_list <- list(
  late_fall_ids = "outputs/007/late-fall-ids.txt",
  cv_fall_ids = "outputs/007/cv-fall-ids.txt",
  candidates_for_assays = "outputs/007/candidates_for_assays.csv"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

@@

Accordingly, we will investigate a 5 Mb chunck of Chromosome 34 to:

@@

### Get the IDs.
```{r, message=FALSE}

@@

ggsave(g, filename = "tex/images/lfar-candidates.pdf", width = 8, height = 5)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

@@

```{r}
# inputs:
input_list <- list(
  chr34_late_fall_mafs = "stored_results/5Mb_Chr34-late-fall.mafs.gz",
  chr34_cv_fall_mafs = "stored_results/5Mb_Chr34-cv-fall.mafs.gz",
  gwas_lf_v_sanjo = "stored_results/lfar-gwas/lf_v_sanjo-NC_037130.1.lrt0.gz",
  gwas_lf_v_frhf = "stored_results/lfar-gwas/lf_v_frhf-NC_037130.1.lrt0.gz",
  ots34_snpeff = "stored_results/thompson2020-vcf/ots34-5Mb-snpEff.vcf.gz",
  pvals = "stored_results/stored_p_vals.rds"
)
# outputs:
output_list <- list(
  lfar_candidates_tex = "tex/images/lfar-candidates.pdf"
)

# additional outputs not specified in Snakefile:
output_list <-  c(
  output_list,
  late_fall_ids = "results/lfar/late-fall-ids.txt",
  cv_fall_ids = "results/lfar/cv-fall-ids.txt",
  candidates_for_assays = "results/lfar/candidates_for_assays.csv"
)

# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

@@

Accordingly, we will investigate a 5 Mb chunk of Chromosome 34 to:

@@

### Get the IDs
```{r, message=FALSE}

@@

ggsave(g, filename = output_list$lfar_candidates, width = 8, height = 5)
```
